---
title: "Single-case analysis with PhysioAppKit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single-case analysis with PhysioAppKit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
```

PhysioAppKit is a small, domain-neutral engine for single-case (N-of-1)
analysis. It works on plain numeric vectors and generic "group A vs group B"
contrasts; a domain application layer (rehabilitation, sport,
psychophysiology, ...) supplies the ontology, thresholds and reporting and calls
these functions. This vignette walks the two things the engine does: a
single-case effect-size + verdict, and ordinal-to-interval Rasch measurement.
Everything runs offline on tiny synthetic data.

```{r setup}
library(PhysioAppKit)
```

## A single-case A-vs-B contrast

Suppose a baseline phase (A) of repeated observations is followed by an
intervention phase (B), and a change of `threshold` units is the smallest that
matters (an MCID in rehab, an SWC in sport, a SESOI in psychophysiology).

```{r analyze}
baseline     <- c(10, 11, 9, 10, 10)
intervention <- c(12, 13, 14, 15)

res <- nonoverlap_analyze(baseline, intervention, threshold = 2,
                          direction = "increase")
c(nap = res$nap, interpretation = res$interpretation,
  beyond_threshold = res$beyond_threshold)
```

`nap()` is the Nonoverlap of All Pairs effect size (0.5 = complete overlap / no
effect, 1 = every B point beats every A point), and `interpret_nap()` bands it:

```{r nap}
nap(baseline, intervention)
interpret_nap(res$nap)
```

A phase plot shows the two phases against the baseline decision band and the
threshold line:

```{r plot}
y <- c(baseline, intervention)
phase_plot(x = seq_along(y), y = y,
           a_idx = seq_along(baseline),
           b_idx = length(baseline) + seq_along(intervention),
           center = res$baseline_mean, lo = res$band_lo, hi = res$band_hi,
           thr = res$baseline_mean + res$threshold, thr_lab = "threshold",
           ylab = "outcome", main = "Single-case phase plot",
           a_lab = "Baseline", b_lab = "Intervention")
```

When several outcomes are each judged separately, `combine_verdicts()` reduces
them to one honest overall label (all supported, some supported, or none):

```{r verdicts}
combine_verdicts(list(
  list(supported = res$beyond_threshold),
  list(supported = TRUE),
  list(supported = NA)      # inconclusive
))
```

## Ordinal ratings to an interval scale (Rasch)

A raw sum of ordinal items is not an interval scale: equal raw gains mean
different real gains at different score levels. The dichotomous Rasch model
calibrates item difficulties and places each person on a linear (logit) scale.

```{r rasch}
set.seed(1)
ability    <- rnorm(40)
difficulty <- c(-1.5, -0.5, 0, 0.5, 1.5)
responses  <- sapply(difficulty, function(d) rbinom(40, 1, plogis(ability - d)))

fit <- rasch_measure(responses)
round(fit$delta, 2)        # calibrated item difficulties (centred logits)
```

The raw-score-to-measure table exposes the non-linear ogive --- equal raw-score
steps are unequal logit steps, largest near the floor and ceiling:

```{r ogive}
raw_score_measure(fit$delta)
```

For ordered-category (polytomous) scales --- ICF qualifiers, Likert, the Barthel
Index --- use `pcm_measure()` (Partial Credit / Rating Scale models) and
`poly_raw_measure()`; `pcm_stack()` calibrates repeated measures onto one common
scale. See `?PhysioAppKit` for the full map.
